---
title: "5205 Group Project"
author: "Group 8"
output: html_notebook
---

We would like to understand whether covid is a factor makes the unemployment rate fluctuates during 2020 and 2021. First of all, we have to make sure the data is started from 2020, because the case number of covid boosted from 2020 March.
```{r}
data <- read.csv('final_data.csv')

covid <- data[-c(1:10),] 
covid <- covid[-c(3:12),] 
covid <- covid[-c(5:14),] 
covid <- covid[-c(7:16),] 
covid <- covid[-c(9:18),] 
covid <- covid[-c(11:20),] 
covid <- covid[-c(13:22),] 
covid <- covid[-c(15:24),] 
covid <- covid[-c(17:26),] 
covid <- covid[-c(19:28),]
covid <- covid[-c(21:30),] 
covid <- covid[-c(23:32),] 
```


Linear regression is used to test the relationship between covid case and unemployment rate. A p value smaller than 0.05 is considered as significant, which indicates a linear relationship between those two factors. 

US case and educational level
```{r}
##Primary_School
model_ps <- lm(Primary_School~US_case, covid)
summary(model_ps)
```

```{r}
##High_School
model_hs <- lm(High_School~US_case, covid)
summary(model_hs)
```

```{r}
##Associates_Degree
model_ad <- lm(Associates_Degree~US_case, covid)
summary(model_ad)
```

```{r}
##Professional_Degree
model_pd <- lm(Professional_Degree~US_case, covid)
summary(model_pd)
```

US case and race
```{r}
##White
model_w <- lm(White~US_case, covid)
summary(model_w)
```

```{r}
##Black
model_b <- lm(Black~US_case, covid)
summary(model_b)
```

```{r}
##Asian
model_a <- lm(Asian~US_case, covid)
summary(model_a)
```

```{r}
##Hispanic
model_h <- lm(Hispanic~US_case, covid)
summary(model_h)
```

US case and gender
```{r}
##Men
model_men <- lm(Men~US_case, covid)
summary(model_men)
```

```{r}
##Women
model_women <- lm(Women~US_case, covid)
summary(model_women)
```
All of the p values are less than 0.05. Therefore, we can conclude that covid does influence US unemployment rate of different people.

To make further interpretation, we select the top 4 populous states in the US to test the influence of covid on unemployment rate by linear regression.

```{r}
##California
model_cal <- lm(California~Cal_case, covid)
summary(model_cal)
```

```{r}
##Florida
model_flo <- lm(Florida~Flo_case, covid)
summary(model_flo)
```

```{r}
##Texas
model_tex <- lm(Texas~Tex_case, covid)
summary(model_tex)
```

```{r}
##New York
model_ny <- lm(NY~NY_case, covid)
summary(model_ny)
```
Similarly, covid does impact on individual state's unemployment rate, especially Florida.



```{r}
unemp <- read.csv('final_data.csv')
str(unemp)
head(unemp,10)
names(unemp)
plot(unemp[2:12,3])
hist(unemp$Primary_School)
hist(unemp$High_School)
```
Here I imported the data and I have found that the original data set put all the Januaries together, Febuaries together, etc, not the way I wanted. So I regrouped the data, making it in rank of years.



```{r}
library('ggplot2')
library('forecast')
library('tseries')
```



REGROUP CODE: 
```{r}
unemp10<-unemp[unemp$Year=="2010",]
unemp11<-unemp[unemp$Year=="2011",]
unemp12<-unemp[unemp$Year=="2012",]
unemp13<-unemp[unemp$Year=="2013",]
unemp14<-unemp[unemp$Year=="2014",]
unemp15<-unemp[unemp$Year=="2015",]
unemp16<-unemp[unemp$Year=="2016",]
unemp17<-unemp[unemp$Year=="2017",]
unemp18<-unemp[unemp$Year=="2018",]
unemp19<-unemp[unemp$Year=="2019",]
unemp20<-unemp[unemp$Year=="2020",]
unemp21<-unemp[unemp$Year=="2021",]
unempall.1 <- rbind(unemp10, unemp11)
unempall.2<- rbind(unempall.1, unemp12)
unempall.3<- rbind(unempall.2, unemp13)
unempall.4<- rbind(unempall.3, unemp14)
unempall.5<- rbind(unempall.4, unemp15)
unempall.6<- rbind(unempall.5, unemp16)
unempall.7<- rbind(unempall.6, unemp17)
unempall.8<- rbind(unempall.7, unemp18)
unempall.9<- rbind(unempall.8, unemp19)
unempall.10<- rbind(unempall.9, unemp20)
unempall<- rbind(unempall.10, unemp21)
```


Split the data into train and test
（we delete two months from test group due to length match in later accuracy function）
```{r}
train = unempall.10
test = unemp21[-c(11,12),]

train
test
```



The first graph is grouped by races
```{r}
unempall
plot(unempall$White, type='l', xaxt='n',ylim=c(0,28), ylab= "unemployment rate")
legend(75, 28, legend=c("White", "Asian", "Hispanic", "Black"),
       col=c("black", "red", "blue", "pink"), lty=1:2, cex=0.8)
axis(1, labels=2010:2022, at=(0:12)*12+1)
points(unempall$Asian,type='l', xaxt='n', col = "red")
points(unempall$Hispanic,type='l', xaxt='n', col = "blue")
points(unempall$Black,type='l', xaxt='n', col = "pink")
```



The second graph is grouped by academic background
```{r}
plot(unempall$Primary_School, type='l', xaxt='n',ylim=c(0,28), ylab= "unemployment rate")
legend(55, 28, legend=c("Primary_School", "Highschool", "Associate_Degree", "Professional_Degree"),
       col=c("black", "red", "blue", "pink"), lty=1:2, cex=0.8)
axis(1, labels=2010:2022, at=(0:12)*12+1)
points(unempall$High_School,type='l', xaxt='n', col = "red")
points(unempall$Associates_Degree,type='l', xaxt='n', col = "blue")
points(unempall$Professional_Degree,type='l', xaxt='n', col = "pink")
```



The third graph is grouped by genders
```{r}
plot(unempall$Men, type='l', xaxt='n',ylim=c(0,20), ylab= "unemployment rate")
legend(80, 20, legend=c("Men", "Women"),
       col=c("black", "red"), lty=1:2, cex=0.8)
axis(1, labels=2010:2022, at=(0:12)*12+1)
points(unempall$Women,type='l', xaxt='n', col = "red")
```

The fourth graph is grouped by states.
```{r}
plot(unempall$California, type='l', xaxt='n',ylim=c(0,20), ylab= "unemployment rate")
legend(80, 20, legend=c("California", "Florida", "New_York", "Texas"),
       col=c("black", "red", "blue", "pink"), lty=1:2, cex=0.8)
axis(1, labels=2010:2022, at=(0:12)*12+1)
points(unempall$Florida,type='l', xaxt='n', col = "red")
points(unempall$New_York,type='l', xaxt='n', col = "blue")
points(unempall$Texas,type='l', xaxt='n', col = "pink")
```


This is the arima for Primary school.
```{r}
tsallp <- ts(data=train$Primary_School)
autoarimatsallp <- auto.arima(tsallp,stepwise = F,approximation = F)
forecasttsallp <- forecast(autoarimatsallp)
forecasttsallp$residuals
```

```{r}
plot(forecasttsallp)
```

```{r}
summary(autoarimatsallp)
```

```{r}
length(forecasttsallp)
tsallp_test = test$Primary_School
accuracy(forecasttsallp,tsallp_test)
```




This is the arima for High school.
```{r}
tsallh <- ts(data=train$High_School)
autoarimatsallh <- auto.arima(tsallh)
forecasttsallh <- forecast(autoarimatsallh)
forecasttsallh$residuals
```


```{r}
plot(forecasttsallh)
```

```{r}
summary(autoarimatsallh)
```

```{r}
tsallh_test = test$High_School
tsallh_test
accuracy(forecasttsallh,tsallh_test)
```




This is the arima for Associate degree.
```{r}
tsallASS <- ts(data=train$Associates_Degree)
autoarimatsallASS <- auto.arima(tsallASS)
forecasttsallASS <- forecast(autoarimatsallASS)
forecasttsallASS$residuals
```


```{r}
plot(forecasttsallASS)
```


```{r}
summary(autoarimatsallASS)
```

```{r}
tsallass_test = test$Associates_Degree
tsallass_test
accuracy(forecasttsallASS,tsallass_test)
```



This is the arima for Professional Degree. 
```{r}
tsallPro <- ts(data=train$Professional_Degree)
autoarimatsallPro <- auto.arima(tsallPro)
forecasttsallPro <- forecast(autoarimatsallPro)
forecasttsallPro$residuals
```

```{r}
plot(forecasttsallPro)
```


```{r}
summary(autoarimatsallPro)
```
```{r}
tsallpro_test = test$Professional_Degree
tsallpro_test
accuracy(forecasttsallPro,tsallpro_test)
```



This is the arima for White people.
```{r}
tsallw <- ts(data=train$White)
autoarimatsallw <- auto.arima(tsallw)
forecasttsallw <- forecast(autoarimatsallw)
forecasttsallw$residuals
```

```{r}
plot(forecasttsallw)
```

```{r}
summary(autoarimatsallw)
```

```{r}
tsallw_test = test$White
tsallw_test
accuracy(forecasttsallw,tsallw_test)
```



This is the arima for Asian.
```{r}
tsallA <- ts(data=train$Asian)
autoarimatsallA <- auto.arima(tsallA)
forecasttsallA <- forecast(autoarimatsallA)
forecasttsallA$residuals
```


```{r}
plot(forecasttsallA)
```

```{r}
summary(autoarimatsallA)
```
```{r}
tsalla_test = test$Asian
tsalla_test
accuracy(forecasttsallA,tsalla_test)
```




This is arima for Black people. 
```{r}
tsallB <- ts(data=train$Black)
autoarimatsallB <- auto.arima(tsallB)
forecasttsallB <- forecast(autoarimatsallB)
forecasttsallB$residuals
```

```{r}
plot(forecasttsallB)
```

```{r}
summary(autoarimatsallB)
```
```{r}
tsallb_test = test$Black
tsallb_test
accuracy(forecasttsallB,tsallb_test)
```



This is arima for Hispanic.
```{r}
tsallH <- ts(data=train$Hispanic)
autoarimatsallH <- auto.arima(tsallH)
forecasttsallH <- forecast(autoarimatsallH)
forecasttsallH$residuals
```

```{r}
plot(forecasttsallH)
```

```{r}
summary(autoarimatsallH)
```
```{r}
tsallh_test = test$Hispanic
tsallh_test
accuracy(forecasttsallH,tsallh_test)
```




This is arima for Men.
```{r}
tsallM <- ts(data=train$Men)
autoarimatsallM <- auto.arima(tsallM)
forecasttsallM <- forecast(autoarimatsallM)
forecasttsallM$residuals
```

```{r}
plot(forecasttsallM)
```

```{r}
summary(autoarimatsallM)
```
```{r}
tsallm_test = test$Men
tsallm_test
accuracy(forecasttsallM,tsallm_test)
```




This is arima for Women. 
```{r}
tsallwo <- ts(data=train$Women)
autoarimatsallwo <- auto.arima(tsallwo)
forecasttsallwo <- forecast(autoarimatsallwo)
forecasttsallwo$residuals
```

```{r}
plot(forecasttsallwo)
```

```{r}
summary(autoarimatsallwo)
```
```{r}
tsallwo_test = test$Women
tsallwo_test
accuracy(forecasttsallwo,tsallwo_test)
```




This is arima for CA. 
```{r}
tsallCA <- ts(data=train$California)
autoarimatsallCA <- auto.arima(tsallCA)
forecasttsallCA <- forecast(autoarimatsallCA)
forecasttsallCA$residuals
```

```{r}
plot(forecasttsallCA)
```

```{r}
summary(autoarimatsallCA)
```
```{r}
tsallCA_test = test$California
tsallCA_test
accuracy(forecasttsallCA,tsallCA_test)
```



This is arima for Florida. 
```{r}
tsallF <- ts(data=train$Florida)
autoarimatsallF <- auto.arima(tsallF)
forecasttsallF <- forecast(autoarimatsallF)
forecasttsallF$residuals
```

```{r}
plot(forecasttsallF)
```

```{r}
summary(autoarimatsallF)
```
```{r}
tsallFL_test = test$Florida
tsallFL_test
accuracy(forecasttsallF,tsallFL_test)
```



This is arima for Texas. 
```{r}
tsallT <- ts(data=train$Texas)
autoarimatsallT <- auto.arima(tsallT)
forecasttsallT <- forecast(autoarimatsallT)
forecasttsallT$residuals
```

```{r}
plot(forecasttsallT)
```

```{r}
summary(autoarimatsallT)
```
```{r}
tsallTX_test = test$Texas
tsallTX_test
accuracy(forecasttsallT,tsallTX_test)
```



This is arima for New York. 
```{r}
tsallNY <- ts(data=train$NY)
autoarimatsallNY <- auto.arima(tsallNY)
forecasttsallNY <- forecast(autoarimatsallNY)
forecasttsallNY$residuals
```

```{r}
plot(forecasttsallNY)
```

```{r}
summary(autoarimatsallwo)
```

```{r}
tsallNY_test = test$NY
tsallNY_test
accuracy(forecasttsallNY,tsallNY_test)
```




